---
title: "Network Meta-Analysis"
output: html_notebook
---

Principal Investigator: Aakash Garg  
Statistician: Davit Sargsyan  
Date: `r date()`  
  
```{r setup}
require(netmeta)
require(data.table)
require(DT)
```

# Load data
```{r data}
dt1 <- fread("Short DAPT Excel Sheet New.csv")
datatable(dt1)
```

Next, we separated each outcome, computed log odds ratios and their standard errors, and performed network meta-analysis.  
  
# All-Cause Mortality (ACM)
```{r acm}
tmp <- dt1[, c("Study",
               "Arm",
               "N",
               "ACM")]
names(tmp)[4] <- "Events"
tmp <- tmp[!is.na(Events),]
tmp$Arm <- factor(tmp$Arm,
                  levels = c("DAPT",
                             "ASA only",
                             "P2Y12 only"))

tmp1 <- tmp[Arm != "DAPT",]

tmp2 <-tmp[Arm == "DAPT",
           c("Study",
             "N",
             "Events")] 

tmp3 <- merge(tmp1,
              tmp2,
              by = "Study")

tmp3[, logOR := log((Events.x/N.x)/(Events.y/N.y))]
tmp3[, seLogOR := sqrt((1/Events.x) + 
                         (1/N.x) +
                         (1/Events.y) + 
                         (1/N.y))]
tmp3$Compare <- "DAPT"
datatable(tmp3)

m1 <- netmeta(TE = logOR,
              seTE = seLogOR,
              treat1 = Arm,
              treat2 = Compare,
              studlab = Study,
              reference.group = "DAPT",
              comb.random = TRUE,
              all.treatments = TRUE,
              sm = "OR",
              data = tmp3)
netgraph(m1,
         cex.number.of.studies = TRUE)

print(m1,
      reference.group = "DAPT")
print(m1,
      reference.group = "ASA only")

forest(m1, 
       reference.group = "DAPT",
       col.square = "blue",
       sortvar = c("P2Y12 only",
                   "ASA only",
                   "DAPT"),
       smlab = "")
```

# CV Death
```{r cv_death}
tmp <- dt1[, c("Study",
               "Arm",
               "N",
               "CV Death")]
names(tmp)[4] <- "Events"
tmp <- tmp[!is.na(Events),]
tmp$Arm <- factor(tmp$Arm,
                  levels = c("DAPT",
                             "ASA only",
                             "P2Y12 only"))

tmp1 <- tmp[Arm != "DAPT",]

tmp2 <-tmp[Arm == "DAPT",
           c("Study",
             "N",
             "Events")] 

tmp3 <- merge(tmp1,
              tmp2,
              by = "Study")

tmp3[, logOR := log((Events.x/N.x)/(Events.y/N.y))]
tmp3[, seLogOR := sqrt((1/Events.x) + 
                      (1/N.x) +
                      (1/Events.y) + 
                      (1/N.y))]
tmp3$Compare <- "DAPT"
datatable(tmp3)

m1 <- netmeta(TE = logOR,
              seTE = seLogOR,
              treat1 = Arm,
              treat2 = Compare,
              studlab = Study,
              reference.group = "DAPT",
              comb.random = TRUE,
              all.treatments = TRUE,
              sm = "OR",
              data = tmp3)
print(m1,
      reference.group = "DAPT")
print(m1,
      reference.group = "ASA only")

forest(m1, 
       reference.group = "DAPT",
       col.square = "blue",
       sortvar = c("P2Y12 only",
                   "ASA only",
                   "DAPT"),
       smlab = "")
```

# MI
```{r mi}
tmp <- dt1[, c("Study",
               "Arm",
               "N",
               "MI")]
names(tmp)[4] <- "Events"
tmp <- tmp[!is.na(Events),]
tmp$Arm <- factor(tmp$Arm,
                  levels = c("DAPT",
                             "ASA only",
                             "P2Y12 only"))

tmp1 <- tmp[Arm != "DAPT",]

tmp2 <-tmp[Arm == "DAPT",
           c("Study",
             "N",
             "Events")] 

tmp3 <- merge(tmp1,
              tmp2,
              by = "Study")

tmp3[, logOR := log((Events.x/N.x)/(Events.y/N.y))]
tmp3[, seLogOR := sqrt((1/Events.x) + 
                      (1/N.x) +
                      (1/Events.y) + 
                      (1/N.y))]
tmp3$Compare <- "DAPT"
datatable(tmp3)

m1 <- netmeta(TE = logOR,
              seTE = seLogOR,
              treat1 = Arm,
              treat2 = Compare,
              studlab = Study,
              reference.group = "DAPT",
              comb.random = TRUE,
              all.treatments = TRUE,
              sm = "OR",
              data = tmp3)
print(m1,
      reference.group = "DAPT")
print(m1,
      reference.group = "ASA only")

forest(m1, 
       reference.group = "DAPT",
       col.square = "blue",
       sortvar = c("P2Y12 only",
                   "ASA only",
                   "DAPT"),
       smlab = "")
```

# Stroke
```{r stroke}
tmp <- dt1[, c("Study",
               "Arm",
               "N",
               "Stroke")]
names(tmp)[4] <- "Events"
tmp <- tmp[!is.na(Events),]
tmp$Arm <- factor(tmp$Arm,
                  levels = c("DAPT",
                             "ASA only",
                             "P2Y12 only"))

tmp1 <- tmp[Arm != "DAPT",]

tmp2 <-tmp[Arm == "DAPT",
           c("Study",
             "N",
             "Events")] 

tmp3 <- merge(tmp1,
              tmp2,
              by = "Study")

tmp3[, logOR := log((Events.x/N.x)/(Events.y/N.y))]
tmp3[, seLogOR := sqrt((1/Events.x) + 
                      (1/N.x) +
                      (1/Events.y) + 
                      (1/N.y))]
tmp3$Compare <- "DAPT"
datatable(tmp3)

m1 <- netmeta(TE = logOR,
              seTE = seLogOR,
              treat1 = Arm,
              treat2 = Compare,
              studlab = Study,
              reference.group = "DAPT",
              comb.random = TRUE,
              all.treatments = TRUE,
              sm = "OR",
              data = tmp3)
print(m1,
      reference.group = "DAPT")
print(m1,
      reference.group = "ASA only")

forest(m1, 
       reference.group = "DAPT",
       col.square = "blue",
       sortvar = c("P2Y12 only",
                   "ASA only",
                   "DAPT"),
       smlab = "")
```

# ST
```{r st}
tmp <- dt1[, c("Study",
               "Arm",
               "N",
               "ST")]
names(tmp)[4] <- "Events"
tmp <- tmp[!is.na(Events),]
tmp$Arm <- factor(tmp$Arm,
                  levels = c("DAPT",
                             "ASA only",
                             "P2Y12 only"))

tmp1 <- tmp[Arm != "DAPT",]

tmp2 <-tmp[Arm == "DAPT",
           c("Study",
             "N",
             "Events")] 

tmp3 <- merge(tmp1,
              tmp2,
              by = "Study")

tmp3[, logOR := log((Events.x/N.x)/(Events.y/N.y))]
tmp3[, seLogOR := sqrt((1/Events.x) + 
                      (1/N.x) +
                      (1/Events.y) + 
                      (1/N.y))]
tmp3$Compare <- "DAPT"
datatable(tmp3)

m1 <- netmeta(TE = logOR,
              seTE = seLogOR,
              treat1 = Arm,
              treat2 = Compare,
              studlab = Study,
              reference.group = "DAPT",
              comb.random = TRUE,
              all.treatments = TRUE,
              sm = "OR",
              data = tmp3)
print(m1,
      reference.group = "DAPT")
print(m1,
      reference.group = "ASA only")

forest(m1, 
       reference.group = "DAPT",
       col.square = "blue",
       sortvar = c("P2Y12 only",
                   "ASA only",
                   "DAPT"),
       smlab = "")
```

# Major Bleeding
```{r major_bleed}
tmp <- dt1[, c("Study",
               "Arm",
               "N",
               "Major Bleeding")]
names(tmp)[4] <- "Events"
tmp <- tmp[!is.na(Events),]
tmp$Arm <- factor(tmp$Arm,
                  levels = c("DAPT",
                             "ASA only",
                             "P2Y12 only"))

tmp1 <- tmp[Arm != "DAPT",]

tmp2 <-tmp[Arm == "DAPT",
           c("Study",
             "N",
             "Events")] 

tmp3 <- merge(tmp1,
              tmp2,
              by = "Study")

tmp3[, logOR := log((Events.x/N.x)/(Events.y/N.y))]
tmp3[, seLogOR := sqrt((1/Events.x) + 
                         (1/N.x) +
                         (1/Events.y) + 
                         (1/N.y))]
tmp3$Compare <- "DAPT"
datatable(tmp3)

m1 <- netmeta(TE = logOR,
              seTE = seLogOR,
              treat1 = Arm,
              treat2 = Compare,
              studlab = Study,
              reference.group = "DAPT",
              comb.random = TRUE,
              all.treatments = TRUE,
              sm = "OR",
              data = tmp3)
print(m1,
      reference.group = "DAPT")
print(m1,
      reference.group = "ASA only")

forest(m1, 
       reference.group = "DAPT",
       col.square = "blue",
       sortvar = c("P2Y12 only",
                   "ASA only",
                   "DAPT"),
       smlab = "")
```

# Major or Minor Bleeding
```{r major_and_minor_bleed}
tmp <- dt1[, c("Study",
               "Arm",
               "N",
               "Major or Minor Bleeding")]
names(tmp)[4] <- "Events"
tmp <- tmp[!is.na(Events),]
tmp$Arm <- factor(tmp$Arm,
                  levels = c("DAPT",
                             "ASA only",
                             "P2Y12 only"))

tmp1 <- tmp[Arm != "DAPT",]

tmp2 <-tmp[Arm == "DAPT",
           c("Study",
             "N",
             "Events")] 

tmp3 <- merge(tmp1,
              tmp2,
              by = "Study")

tmp3[, logOR := log((Events.x/N.x)/(Events.y/N.y))]
tmp3[, seLogOR := sqrt((1/Events.x) + 
                      (1/N.x) +
                      (1/Events.y) + 
                      (1/N.y))]
tmp3$Compare <- "DAPT"
datatable(tmp3)

m1 <- netmeta(TE = logOR,
              seTE = seLogOR,
              treat1 = Arm,
              treat2 = Compare,
              studlab = Study,
              reference.group = "DAPT",
              comb.random = TRUE,
              all.treatments = TRUE,
              sm = "OR",
              data = tmp3)
print(m1,
      reference.group = "DAPT")
print(m1,
      reference.group = "ASA only")

forest(m1, 
       reference.group = "DAPT",
       col.square = "blue",
       sortvar = c("P2Y12 only",
                   "ASA only",
                   "DAPT"),
       smlab = "")
```

# CV Death or MI 
```{r cv_death_and_mi}
tmp <- dt1[, c("Study",
               "Arm",
               "N",
               "CV Death or MI")]
names(tmp)[4] <- "Events"
tmp <- tmp[!is.na(Events),]
tmp$Arm <- factor(tmp$Arm,
                  levels = c("DAPT",
                             "ASA only",
                             "P2Y12 only"))

tmp1 <- tmp[Arm != "DAPT",]

tmp2 <-tmp[Arm == "DAPT",
           c("Study",
             "N",
             "Events")] 

tmp3 <- merge(tmp1,
              tmp2,
              by = "Study")

tmp3[, logOR := log((Events.x/N.x)/(Events.y/N.y))]
tmp3[, seLogOR := sqrt((1/Events.x) + 
                      (1/N.x) +
                      (1/Events.y) + 
                      (1/N.y))]
tmp3$Compare <- "DAPT"
datatable(tmp3)

m1 <- netmeta(TE = logOR,
              seTE = seLogOR,
              treat1 = Arm,
              treat2 = Compare,
              studlab = Study,
              reference.group = "DAPT",
              comb.random = TRUE,
              all.treatments = TRUE,
              sm = "OR",
              data = tmp3)
print(m1,
      reference.group = "DAPT")
print(m1,
      reference.group = "ASA only")

forest(m1, 
       reference.group = "DAPT",
       col.square = "blue",
       sortvar = c("P2Y12 only",
                   "ASA only",
                   "DAPT"),
       smlab = "")
```

# Composite of Ischemic Events 
```{r composite_ischemic}
tmp <- dt1[, c("Study",
               "Arm",
               "N",
               "Composite of Ischemic Events")]
names(tmp)[4] <- "Events"
tmp <- tmp[!is.na(Events),]
tmp$Arm <- factor(tmp$Arm,
                  levels = c("DAPT",
                             "ASA only",
                             "P2Y12 only"))

tmp1 <- tmp[Arm != "DAPT",]

tmp2 <-tmp[Arm == "DAPT",
           c("Study",
             "N",
             "Events")] 

tmp3 <- merge(tmp1,
              tmp2,
              by = "Study")

tmp3[, logOR := log((Events.x/N.x)/(Events.y/N.y))]
tmp3[, seLogOR := sqrt((1/Events.x) + 
                      (1/N.x) +
                      (1/Events.y) + 
                      (1/N.y))]
tmp3$Compare <- "DAPT"
datatable(tmp3)

m1 <- netmeta(TE = logOR,
              seTE = seLogOR,
              treat1 = Arm,
              treat2 = Compare,
              studlab = Study,
              reference.group = "DAPT",
              comb.random = TRUE,
              all.treatments = TRUE,
              sm = "OR",
              data = tmp3)
print(m1,
      reference.group = "DAPT")
print(m1,
      reference.group = "ASA only")

forest(m1, 
       reference.group = "DAPT",
       col.square = "blue",
       sortvar = c("P2Y12 only",
                   "ASA only",
                   "DAPT"),
       smlab = "")
```

# References
1. [Doing Meta-Analysis in R Meta-Analysis, A Hands-On Guide](https://bookdown.org/MathiasHarrer/Doing_Meta_Analysis_in_R/frequentist-network-meta-analysis.html), Chapter 11.1 Frequentist Network, M. Harrer et al.  
2. [An Introduction to Categorical data Analysis](https://mregresion.files.wordpress.com/2012/08/agresti-introduction-to-categorical-data.pdf), Alan Agresti, 2007, Page 30.  
3. [Safety and Efficacy of Antithrombotic Strategies in Patients With Atrial Fibrillation Undergoing Percutaneous Coronary Intervention
A Network Meta-analysis of Randomized Controlled Trials](https://jamanetwork.com/journals/jamacardiology/article-abstract/2735647), R. D. Lopes et al, Jama Cardiology, June 19 2019  
4. [Mortality in patients treated with extended duration dual antiplatelet therapy after drug-eluting stent implantation: a pairwise and Bayesian network meta-analysis of randomised trials](https://www.sciencedirect.com/science/article/pii/S014067361560263X?via%3Dihub), T. Palmerini, The Lancet, June 13 2015  
  
# Session Information
```{r session}
sessionInfo()
```