---
title: "Network Meta-Analysis, Version 6"
output:
  html_notebook:
    number_sections: yes
    toc: TRUE
    toc_float: TRUE
    code_folding: hide
---

Principal Investigator: Aakash Garg  
Statistician: Davit Sargsyan  
Date: `r date()`  
  
```{r setup, include = FALSE}
require(netmeta)
require(data.table)
require(DT)
require(ggplot2)
require(scales)
```

```{r function_3groups}
foo3 <- function(ep) {
  tmp <- dt1[, c("Study",
                 "Arm",
                 "N",
                 ep),
             with = FALSE]
  names(tmp)[4] <- "Events"
  tmp <- tmp[!is.na(Events),]
  tmp$Arm <- factor(tmp$Arm,
                    levels = c("DAPT",
                               "ASA only",
                               "P2Y12 only"))
  
  tmp1 <- tmp[Arm != "DAPT",]
  
  tmp2 <- tmp[Arm == "DAPT",
              c("Study",
                "N",
                "Events")] 
  
  tmp3 <- merge(tmp1,
                tmp2,
                by = "Study")
  
  # Add half to all counts to avoid Infs
  tmp3.1 <- copy(tmp3)
  tmp3.1$N.x <- tmp3.1$N.x + 0.5
  tmp3.1$Events.x <- tmp3.1$Events.x + 0.5
  tmp3.1$N.y <- tmp3.1$N.y + 0.5
  tmp3.1$Events.y <- tmp3.1$Events.y + 0.5
  
  tmp3.1[, OR := (Events.x/N.x)/(Events.y/N.y)]
  tmp3.1[, logOR := log(OR)]
  tmp3.1[, seLogOR := sqrt((1/Events.x) + 
                             (1/N.x) +
                             (1/Events.y) + 
                             (1/N.y))]
  tmp3.1$Compare <- "DAPT"
  tmp3.1 <- tmp3.1[, c(1:2, 10, 3:9)]
  print(datatable(tmp3.1,
                  caption = "",
                  rownames = FALSE,
                  class = "cell-border stripe",
                  options = list(searching = TRUE,
                                 pageLength = nrow(dt1))) %>%
          formatCurrency(columns = 4:7, 
                         currency = "",
                         digits = 1,
                         mark = ",") %>%
          formatCurrency(columns = 8:10, 
                         currency = "",
                         digits = 3,
                         mark = ","))
  
  # # Exclude infinities
  # tmp3 <- tmp3[is.finite(logOR), ]
  
  m1 <- netmeta(TE = logOR,
                seTE = seLogOR,
                n1 = N.x,
                n2 = N.y,
                event1 = Events.x,
                event2 = Events.y,
                treat1 = Arm,
                treat2 = Compare,
                studlab = Study,
                reference.group = "DAPT",
                seq = c("ASA only",
                        "P2Y12 only",
                        "DAPT"),
                comb.random = TRUE,
                comb.fixed = FALSE,
                all.treatments = TRUE,
                sm = "OR",
                data = tmp3.1)
  
  netgraph(m1,
           cex.number.of.studies = TRUE)
  
  print(m1,
        reference.group = "DAPT")
  # print(m1,
  #       reference.group = "ASA only")
  
  forest(m1, 
         reference.group = "DAPT",
         col.square = "blue",
         sortvar = c("P2Y12 only",
                     "ASA only",
                     "DAPT"),
         smlab = "")
  
  t1 <- melt(m1$TE.random)
  names(t1) <- c("Treatment1",
                 "Treatment2",
                 "OR")
  t1$LB <- melt(m1$lower.random)$value
  t1$UB <- melt(m1$upper.random)$value
  t1$pval <- melt(m1$pval.random)$value
  t1 <- data.table(Endpoint = ep,
                   t1)
  t1 <- t1[t1$OR != 0, ]
  t1$OR <- exp(t1$OR)
  t1$LB <- exp(t1$LB)
  t1$UB <- exp(t1$UB)
  
  print(datatable(t1,
                  caption = "",
                  rownames = FALSE,
                  class = "cell-border stripe",
                  options = list(searching = TRUE,
                                 pageLength = nrow(t1))) %>%
          formatCurrency(columns = 4:7, 
                         currency = "",
                         digits = 3))
  
  t2 <- data.table(Endpoint = ep,
                   Study = m1$studlab,
                   Treatment1 = m1$treat1,
                   Treatment2 = m1$treat2,
                   OR = exp(m1$TE.nma.random),
                   LB = exp(m1$lower.nma.random),
                   UB = exp(m1$upper.nma.random),
                   pVal = m1$pval.nma.random)
  t2.1 <- data.table(Endpoint = ep,
                     Study = t2$Study,
                     Treatment1 = t2$Treatment2,
                     Treatment2 = t2$Treatment1,
                     OR = 1/t2$OR,
                     LB = 1/t2$UB,
                     UB = 1/t2$LB,
                     pVal = t2$pVal)
  t2 <- rbindlist(list(t2,
                       t2.1))
  setorder(t2,
           Treatment2)
  
  print(datatable(t2,
                  caption = "",
                  rownames = FALSE,
                  class = "cell-border stripe",
                  options = list(searching = TRUE,
                                 pageLength = nrow(t2))) %>%
          formatCurrency(columns = 5:8, 
                         currency = "",
                         digits = 3))
  
  t3 <- data.table(Endpoint = ep,
                   Study = tmp3.1$Study,
                   Treatment1 = tmp3.1$Arm,
                   Treatment2 = tmp3.1$Compare,
                   N = tmp3$N.x + tmp3$N.y, # NOTE: this are numbers before adding 0.5
                   OR = tmp3.1$OR,
                   LB = exp(tmp3.1$logOR - 1.96*tmp3.1$seLogOR),
                   UB = exp(tmp3.1$logOR + 1.96*tmp3.1$seLogOR))
  
  tmp <- t1[(Treatment1 %in% c("ASA only",
                               "P2Y12 only") & 
               Treatment2 == "DAPT"), ]
  t3 <- rbindlist(list(t3,
                       data.table(Endpoint = ep,
                                  Study = rep("Average", nrow(tmp)),
                                  Treatment1 = tmp$Treatment1,
                                  Treatment2 = tmp$Treatment2,
                                  N = NA,
                                  OR = tmp$OR,
                                  LB = tmp$LB,
                                  UB = tmp$UB)))
  setorder(t3,
           Treatment1)
  print(datatable(t3,
                  caption = "",
                  rownames = FALSE,
                  class = "cell-border stripe",
                  options = list(searching = TRUE,
                                 pageLength = nrow(t3))) %>%
          formatCurrency(columns = 6:8, 
                         currency = "",
                         digits = 3))
  
  t3$Comparison <- paste(t3$Treatment1,
                         "vs.",
                         t3$Treatment2)
  
  p1 <- ggplot(t3,
               aes(x = Study,
                   y = OR,
                   ymin = LB,
                   ymax = UB,
                   fill = Comparison)) +
    facet_wrap(~ Comparison) +
    coord_flip() +
    geom_errorbar(width = 0.3) +
    geom_hline(yintercept = 1,
               linetype = 2) +
    geom_point(size = 3,
               shape = 23) +
    scale_y_continuous(trans = log_trans(),
                       breaks = c(0.1, 0.2, 0.5, 2^(0:5))) +
    theme(legend.position = "none")
  
  print(p1)
}
```

```{r function_4groups}
foo4 <- function(ep) {
  tmp <- dt1[, c("Study",
                 "Arm2",
                 "N",
                 ep),
             with = FALSE]
  names(tmp)[4] <- "Events"
  tmp <- tmp[!is.na(Events),]
  tmp$Arm2 <- factor(tmp$Arm2,
                     levels = c("DAPT",
                                "ASA only 3m",
                                "ASA only 6m",
                                "P2Y12 only"))
  
  tmp1 <- tmp[Arm2 != "DAPT",]
  
  tmp2 <- tmp[Arm2 == "DAPT",
              c("Study",
                "N",
                "Events")] 
  
  tmp3 <- merge(tmp1,
                tmp2,
                by = "Study")
  
  # Add half to all counts to avoid Infs
  tmp3.1 <- copy(tmp3)
  tmp3.1$N.x <- tmp3.1$N.x + 0.5
  tmp3.1$Events.x <- tmp3.1$Events.x + 0.5
  tmp3.1$N.y <- tmp3.1$N.y + 0.5
  tmp3.1$Events.y <- tmp3.1$Events.y + 0.5
  
  tmp3.1[, OR := (Events.x/N.x)/(Events.y/N.y)]
  tmp3.1[, logOR := log(OR)]
  tmp3.1[, seLogOR := sqrt((1/Events.x) + 
                             (1/N.x) +
                             (1/Events.y) + 
                             (1/N.y))]
  tmp3.1$Compare <- "DAPT"
  tmp3.1 <- tmp3.1[, c(1:2, 10, 3:9)]
  print(datatable(tmp3.1,
                  caption = "",
                  rownames = FALSE,
                  class = "cell-border stripe",
                  options = list(searching = TRUE,
                                 pageLength = nrow(dt1))) %>%
          formatCurrency(columns = 4:7, 
                         currency = "",
                         digits = 1,
                         mark = ",") %>%
          formatCurrency(columns = 8:10, 
                         currency = "",
                         digits = 3,
                         mark = ","))
  
  # # Exclude infinities
  # tmp3 <- tmp3[is.finite(logOR), ]
  
  m1 <- netmeta(TE = logOR,
                seTE = seLogOR,
                n1 = N.x,
                n2 = N.y,
                event1 = Events.x,
                event2 = Events.y,
                treat1 = Arm2,
                treat2 = Compare,
                studlab = Study,
                reference.group = "DAPT",
                seq = c("ASA only 3m",
                        "ASA only 6m",
                        "P2Y12 only",
                        "DAPT"),
                comb.random = TRUE,
                comb.fixed = FALSE,
                all.treatments = TRUE,
                sm = "OR",
                data = tmp3.1)
  
  netgraph(m1,
           cex.number.of.studies = TRUE)
  
  print(m1,
        reference.group = "DAPT")
  # print(m1,
  #       reference.group = "ASA only 3m")
  
  forest(m1, 
         reference.group = "DAPT",
         col.square = "blue",
         sortvar = c("P2Y12 only",
                     "ASA only 3m",
                     "ASA only 6m",
                     "DAPT"),
         smlab = "")
  
  t1 <- melt(m1$TE.random)
  names(t1) <- c("Treatment1",
                 "Treatment2",
                 "OR")
  t1$LB <- melt(m1$lower.random)$value
  t1$UB <- melt(m1$upper.random)$value
  t1$pval <- melt(m1$pval.random)$value
  t1 <- data.table(Endpoint = ep,
                   t1)
  t1 <- t1[t1$OR != 0, ]
  t1$OR <- exp(t1$OR)
  t1$LB <- exp(t1$LB)
  t1$UB <- exp(t1$UB)
  
  print(datatable(t1,
                  caption = "",
                  rownames = FALSE,
                  class = "cell-border stripe",
                  options = list(searching = TRUE,
                                 pageLength = nrow(t1))) %>%
          formatCurrency(columns = 4:7, 
                         currency = "",
                         digits = 3))
  
  t2 <- data.table(Endpoint = ep,
                   Study = m1$studlab,
                   Treatment1 = m1$treat1,
                   Treatment2 = m1$treat2,
                   OR = exp(m1$TE.nma.random),
                   LB = exp(m1$lower.nma.random),
                   UB = exp(m1$upper.nma.random),
                   pVal = m1$pval.nma.random)
  t2.1 <- data.table(Endpoint = ep,
                     Study = t2$Study,
                     Treatment1 = t2$Treatment2,
                     Treatment2 = t2$Treatment1,
                     OR = 1/t2$OR,
                     LB = 1/t2$UB,
                     UB = 1/t2$LB,
                     pVal = t2$pVal)
  t2 <- rbindlist(list(t2,
                       t2.1))
  setorder(t2,
           Treatment2)
  
  print(datatable(t2,
                  caption = "",
                  rownames = FALSE,
                  class = "cell-border stripe",
                  options = list(searching = TRUE,
                                 pageLength = nrow(t2))) %>%
          formatCurrency(columns = 5:8, 
                         currency = "",
                         digits = 3))
  
  t3 <- data.table(Endpoint = ep,
                   Study = tmp3.1$Study,
                   Treatment1 = tmp3.1$Arm2,
                   Treatment2 = tmp3.1$Compare,
                   N = tmp3$N.x + tmp3$N.y, # NOTE: this are numbers before adding 0.5
                   OR = tmp3.1$OR,
                   LB = exp(tmp3.1$logOR - 1.96*tmp3.1$seLogOR),
                   UB = exp(tmp3.1$logOR + 1.96*tmp3.1$seLogOR))
  
  tmp <- t1[(Treatment1 %in% c("ASA only 3m",
                               "ASA only 6m",
                               "P2Y12 only") & 
               Treatment2 == "DAPT"), ]
  t3 <- rbindlist(list(t3,
                       data.table(Endpoint = ep,
                                  Study = rep("Average", nrow(tmp)),
                                  Treatment1 = tmp$Treatment1,
                                  Treatment2 = tmp$Treatment2,
                                  N = NA,
                                  OR = tmp$OR,
                                  LB = tmp$LB,
                                  UB = tmp$UB)))
  
  setorder(t3,
           Treatment1)
  print(datatable(t3,
                  caption = "",
                  rownames = FALSE,
                  class = "cell-border stripe",
                  options = list(searching = TRUE,
                                 pageLength = nrow(t3))) %>%
          formatCurrency(columns = 6:8, 
                         currency = "",
                         digits = 3))
  
  t3$Comparison <- paste(t3$Treatment1,
                         "vs.",
                         t3$Treatment2)
  
  p1 <- ggplot(t3,
               aes(x = Study,
                   y = OR,
                   ymin = LB,
                   ymax = UB,
                   fill = Comparison)) +
    facet_wrap(~ Comparison) +
    coord_flip() +
    geom_errorbar(width = 0.3) +
    geom_hline(yintercept = 1,
               linetype = 2) +
    geom_point(size = 3,
               shape = 23) +
    scale_y_continuous(trans = log_trans(),
                       breaks = c(0.1, 0.2, 0.5, 2^(0:5))) +
    theme(legend.position = "none")
  
  print(p1)
}
```

# Load data
```{r data}
dt1 <- fread("data/Short DAPT Excel Sheet 11032019.csv")

# # Global Leaders 1-year follow-up
# dt1 <- dt1[Study != "GLOBAL LEADERS 2Y", ]

# Global Leaders 2-year follow-up
dt1 <- dt1[Study != "GLOBAL LEADERS", ]

datatable(dt1,
          caption = "Table 1: number of patients",
          rownames = FALSE,
          class = "cell-border stripe",
          options = list(searching = TRUE,
                         pageLength = nrow(dt1))) %>%
    formatCurrency(columns = 3:ncol(dt1), 
                 currency = "",
                 digits = 0,
                 mark = ",")
```

```{r network_plot, fig.height = 5, fig.width = 6}
# dt1
# unique(dt1$Study)
# length(unique(dt1$Study))
# unique(dt1$Arm)
# sum(dt1$N)
# 
# length(unique(dt1$Study[dt1$Arm == "ASA only"]))
# length(unique(dt1$Study[dt1$Arm == "P2Y12 only"]))
# length(unique(dt1$Study[dt1$Arm == "DAPT"]))

tmp <- data.table(x = c(1, 2, 3),
                  y = c(1, 2, 1),
                  n = c(8, 10, 2))
p1 <- ggplot() +
  geom_segment(aes(x = c(1, 2, 1),
                   xend = c(2, 3, 3),
                   y = c(1, 2, 1),
                   yend = c(2, 1, 1),
                   size = c(4, 2, 1),
                   linetype = c("dashed",
                                "dashed",
                                "solid")),
               color = "black") +
  geom_segment(aes(x = c(1, 2),
                   xend = c(2, 3),
                   y = c(1, 2),
                   yend = c(2, 1),
                   size = c(2.8, 1.2)),
               color = "white") +
  geom_point(data = tmp,
             aes(x = x,
                 y = y,
                 size = n),
             shape = 21,
             fill = "white") +
  geom_text(aes(x = c(1.2, 2, 2.8),
                y = c(0.9, 2.1, 0.9),
                label = c("S-DAPT/ASA",
                          "L-DAPT",
                          "S-DAPT/P2Y12"))) +
  geom_text(aes(x = c(1.5, 2.5),
                y = c(1.6, 1.6),
                label = c("14 Studies",
                          "4 Studies"),
                angle = c(52, -52))) +
  theme_void() +
  theme(legend.position="none")

tiff(filename = "tmp/network.tiff",
     height = 5,
     width = 6,
     units = 'in',
     res = 600,
     compression = "lzw+p")
print(p1)
graphics.off()

print(p1)
```

Next, we separated each outcome, computed log odds ratios and their standard errors, and performed network meta-analysis.  
  
# All-Cause Mortality (ACM)
## 3 Groups
```{r acm_3, fig.width = 8, fig.height = 4}
foo3("ACM")
```

## 4 Groups
```{r acm_4, fig.width = 12, fig.height = 4}
foo4("ACM")
```

# CV Death
## 3 Groups
```{r cvdeath_3, fig.width = 8, fig.height = 4}
foo3("CV Death")
```

## 4 Groups
```{r cvdeath_4, fig.width = 12, fig.height = 4}
foo4("CV Death")
```

# MI
## 3 Groups
```{r mi_3, fig.width = 8, fig.height = 4}
foo3("MI")
```

## 4 Groups
```{r mi_4, fig.width = 12, fig.height = 4}
foo4("MI")
```

# Stroke
## 3 Groups
```{r stroke_3, fig.width = 8, fig.height = 4}
foo3("Stroke")
```

## 4 Groups
```{r stroke_4, fig.width = 12, fig.height = 4}
foo4("Stroke")
```

# Stent Thrombosis (ST)
## 3 Groups
```{r st_3, fig.width = 8, fig.height = 4}
foo3("ST")
```

## 4 Groups
```{r st_4, fig.width = 12, fig.height = 4}
foo4("ST")
```

# Major Bleeding
## 3 Groups
```{r mb_3, fig.width = 8, fig.height = 4}
foo3("Major Bleeding")
```

## 4 Groups
```{r mb_4, fig.width = 12, fig.height = 4}
foo4("Major Bleeding")
```

# Major or Minor Bleeding
## 3 Groups
```{r mmb_3, fig.width = 8, fig.height = 4}
foo3("Major or Minor Bleeding")
```

## 4 Groups
```{r mmb_4, fig.width = 12, fig.height = 4}
foo4("Major or Minor Bleeding")
```

# Death or MI
## 3 Groups
```{r pep_3, fig.width = 8, fig.height = 4}
foo3("Death or MI")
```

## 4 Groups
```{r pep_4, fig.width = 12, fig.height = 4}
foo4("Death or MI")
```

# Sensitivity analysis: MI only, GLOBAL LEADERS 1-year follow-up
```{r mi_3_no_smart_date, fig.width = 8, fig.height = 4}
# dt1 <- dt1[Study != "SMART DATE", ]
# foo3("MI")
```

# Net Effect: MI - Major Bleeding
```{r net_effect_mi_mbleed, fig.width = 7, fig.height = 5}
dt1$net_mi_majbleed <- dt1$MI - dt1$`Major Bleeding`

p1 <- ggplot(dt1,
             aes(x = net_mi_majbleed,
                 y = Study,
                 fill = Study,
                 size = log(N))) +
  facet_wrap(~ Arm) +
  geom_point(shape = 21) +
  geom_vline(xintercept = 0,
             linetype = "dashed") +
  scale_x_continuous("Net Effect: MI - Major Bleeding") +
  theme(legend.position = "none")
print(p1)

p2 <- ggplot(dt1,
             aes(x = 100*net_mi_majbleed/N,
                 y = Study,
                 fill = Study,
                 size = log(N))) +
  facet_wrap(~ Arm) +
  geom_point(shape = 21) +
  geom_vline(xintercept = 0,
             linetype = "dashed") +
  scale_x_continuous("Net Effect: %MI - %Major Bleeding") +
  theme(legend.position = "none")
print(p2)

dt1[, wgt := N/sum(N),
    by = "Arm"]

dt1[, mu := mean(net_mi_majbleed,
                 na.rm = TRUE),
    by = "Arm"]
dt1[, wgt.mu := sum(wgt*net_mi_majbleed,
                     na.rm = TRUE),
    by = "Arm"]

dt1[, mu.pct := mean(100*net_mi_majbleed/N,
                     na.rm = TRUE),
    by = "Arm"]
dt1[, wgt.mu.pct := sum(100*wgt*net_mi_majbleed/N,
                         na.rm = TRUE),
    by = "Arm"]


t1 <- unique(dt1[, c("Arm",
                     "mu",
                     "wgt.mu",
                     "mu.pct",
                     "wgt.mu.pct")])
colnames(t1) <- c("Arm",
                  "Mean Difference",
                  "Mean Weighted Difference",
                  "Mean % Difference",
                  "Mean Weighted % Difference")
datatable(t1) %>%
  formatCurrency(columns = 2:5,
                 currency = "",
                 digits = 3)
```


# References
1. [Doing Meta-Analysis in R Meta-Analysis, A Hands-On Guide](https://bookdown.org/MathiasHarrer/Doing_Meta_Analysis_in_R/frequentist-network-meta-analysis.html), Chapter 11.1 Frequentist Network, M. Harrer et al.  
2. [An Introduction to Categorical data Analysis](https://mregresion.files.wordpress.com/2012/08/agresti-introduction-to-categorical-data.pdf), Alan Agresti, 2007, Page 30.  
3. [Safety and Efficacy of Antithrombotic Strategies in Patients With Atrial Fibrillation Undergoing Percutaneous Coronary Intervention
A Network Meta-analysis of Randomized Controlled Trials](https://jamanetwork.com/journals/jamacardiology/article-abstract/2735647), R. D. Lopes et al, Jama Cardiology, June 19 2019  
4. [Mortality in patients treated with extended duration dual antiplatelet therapy after drug-eluting stent implantation: a pairwise and Bayesian network meta-analysis of randomised trials](https://www.sciencedirect.com/science/article/pii/S014067361560263X?via%3Dihub), T. Palmerini, The Lancet, June 13 2015  
5. [Meta-Analysis with R](https://www.springer.com/gp/book/9783319214153), Schwarzer, Guido, Carpenter, James R., Rücker, Gerta, Springer 2015  
  
# Abbreviations
```{r abbrev}
abbrev <- data.table(rbind(c("ASA", "Acetylsalicylic acid (aspirin)."),
                           c("DAPT", "Dual antiplatelet therapy (ASA and P2Y12)."),
                           c("P2Y12", "Chemoreceptor for adenosine diphosphate. The drugs clopidogrel (Plavix), prasugrel (Efient, Effient), ticagrelor (Brilinta), and cangrelor (Kengreal) bind to this receptor and are marketed as antiplatelet agents.")))
colnames(abbrev) <- c("Abbreviation",
                      "Description")
setorder(abbrev,
         Abbreviation)
datatable(abbrev)
```

# Session Information
```{r session}
sessionInfo()
```